---
title: "Red Light Therapy Evidence Visualization"
subtitle: "Proposal"
author: 
  - name: "Levi Hood & Vivian Huynh"
    affiliations:
      - name: "College of Information Science, University of Arizona"
description: "Visualizing the current research landscape on red light therapy and mental health outcomes to examine which outcomes are most studied and where the evidence appears strongest, weakest, or most mixed."
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

We will construct our own dataset by collecting and coding published studies on red light therapy (photobiomodulation) and mental health outcomes from sources such as PubMed, review articles, and clinical trial summaries. Each study will be manually reviewed and coded into a structured CSV dataset with the following variables:

-   **Study ID**: Unique identifier
-   **Year**: Publication year
-   **Mental Health Condition**: e.g. depression, anxiety, PTSD, cognitive decline, insomnia, TBI-related mood symptoms
-   **Outcome Category**: Specific health benefit being measured (e.g. mood, memory, sleep quality)
-   **Outcome Direction**: Positive, mixed, or inconclusive
-   **Study Design**: e.g. randomized controlled trial, observational, case study
-   **Application Method**: Transcranial vs. peripheral delivery
-   **Treatment Protocol**: Wavelength, dose, and session duration where reported
-   **Sample Size**: Number of participants
-   **Population**: Clinical vs. healthy sample, age group
-   **Notes**: Brief summary of key findings

A codebook documenting variable definitions and coding decisions will be maintained to ensure transparency and reproducibility. The final dataset will be stored in CSV format and processed in R for cleaning, transformation, and visualization.

## Questions

1.  Which mental health conditions have the most research activity on red light therapy, and how has that activity changed over time?

2.  Where does the evidence appear strongest, most mixed, or most inconclusive — and how does that vary across condition, study design, and application method?

## Analysis Plan

**Question 1 — Research activity by condition over time:**

-   Variables involved: `Mental Health Condition`, `Year`, `Study Design`
-   Visualizations planned:
    -   Bar chart showing number of studies by mental health condition
    -   Timeline plot showing how research activity across conditions has changed over time

**Question 2 — Evidence strength and outcome patterns:**

-   Variables involved: `Outcome Direction`, `Mental Health Condition`, `Study Design`, `Application Method`, `Sample Size`
-   Variables to be created: An aggregated `Evidence Strength` score or category derived from `Outcome Direction` counts per condition, to power the heatmap
-   Visualizations planned:
    -   Stacked bar chart showing outcome direction (positive, mixed, inconclusive) broken down by condition
    -   Heatmap of condition vs. evidence strength, highlighting how much research exists per condition
    -   Dot plot or bubble chart comparing sample sizes across study designs and conditions
    -   Alluvial diagram showing the flow from mental health condition → study design → outcome direction

**Weekly Plan:**

| Week | Task | Status | Assignee | Priority |
|--------------|--------------|--------------|--------------|--------------|
| 1 | Finalize research question and scope | In Progress | Both | High |
| 1 | Identify and collect initial set of studies | In Progress | Both | High |
| 1 | Set up project repository and folder structure | Not Started | Both | Low |
| 2 | Continue study collection and finalize dataset size | Not Started | Both | Low |
| 2 | Develop codebook and variable definitions | Not Started | Vivian | Low |
| 2 | Enter and clean dataset in CSV format | Not Started | Levi | Low |
| 3 | Data cleaning and preprocessing in R | Not Started | Both | Low |
| 3 | Create initial visualizations | Not Started | Both | Low |
| 3 | Review and refine dataset coding consistency | Not Started | Both | Low |
| 4 | Develop visualization dashboard (Shiny) | Not Started | Both | Low |
| 4 | Draft write-up (introduction, methods) | Not Started | Both | Low |
| 4 | Draft visualization explanations | Not Started | Both | Low |
| 5 | Build presentation slides | Not Started | Both | Low |
| 5 | Refine visuals and narrative flow | Not Started | Both | Low |
| 5 | Practice presentation and finalize submission | Not Started | Both | Low |
