---
title: "levi_viz"
output: html_document
---

```{r}

library(plotly)
library(dplyr)
library(ggExtra)

df <- read_csv('/cloud/project/data/tpbm_clean.csv')

```

```{r}
### pie chart
##  population makeup (age)
```


1. Stacked bar chart showing outcome direction (positive, mixed, inconclusive) broken down by condition — note: this would likely replace the simple bar chart above, not complement it, as it contains that information and more
2. Timeline plot showing how research activity across mental health conditions has changed over time

```{r Timeline Plot}

df_years <- df |> group_by(year) |> summarise(total = n())

#ggplot version 
# overtime <- ggplot(df_years,
#        aes(x = year, y = total)) +
#   annotate('rect', xmin = 2017, xmax = 2026, 
#            ymin = -Inf, ymax = Inf, fill = 'darkgreen', alpha = 0.3) +
#   geom_point() +
#   geom_line() +
#   labs(x = 'Year',
#        y = 'Study Count') +
#   annotate('text', label = 'Growth Accelerates', x = 2017, y = 15, angle = 90, hjust = -.2) +
#   annotate('text', label = 'Peak Growth', x = 2025, y = 73, angle = 90, hjust = -.2) +
#   scale_x_continuous(n.breaks = 5)

#plotly version
overtime <- plot_ly(data = df_years,
        x = ~year,
        y = ~total,
        type = 'scatter',
        mode = 'lines+markers') |>
  plotly::layout(xaxis = list(title = 'Year', dtick = 5),
                 yaxis = list(title = 'Study Count'),
                 annotations = list(list(x = 2017,y = 15,text = "Growth Accelerates",showarrow = TRUE,arrowhead = 2),
                                    list(x = 2025,y = 74,text = "Peak Study Count",showarrow = TRUE,arrowhead = 2),
                                    list(x = 2026,y = 5,text = "Still In-Progress",showarrow = TRUE,arrowhead = 2)

))

overtime
```

3. Dot plot or bubble chart comparing sample sizes across study designs and conditions
```{r Scatterplot of Condition vs Sample Size}

p_main <- plot_ly(data = df,
                  x = ~sample,
                  y = ~condition,
                  type = 'scatter',
                  color = ~population_age) |>
  plotly::layout(xaxis = list(title = 'Sample Size', dtick = 20),
                 yaxis = list(title = ''))

p_top <- plot_ly(data = df,
                 x = ~sample,
                 type = 'box')

subplot(p_top, plotly_empty(), p_main, 
        nrows = 2, heights = c(0.1, 0.9), widths = c(0.8, 0.2), 
        shareX = TRUE, shareY = TRUE, margin = 0.01)
```

```{r Table of Study Summary}

```

```{r Tree Map}

### tree map of total samples per condition with annotated poisitive vs negative or so

df_tree <- df |> group_by(condition) |> summarise(total = n())

p_tree <- plot_ly(data = df_tree,
                  type = 'treemap',
                  labels = ~condition,
                  parents = '',
                  values = ~total,
                  marker = list(colors = hcl.colors(nrow(df_tree), palette = "Set1"))
)
p_tree
```


