---
title: "v_viz"
output: html_document
---


```{r}
library(dplyr)
library(forcats)
library(stringr)
library(plotly)
library(tidyverse)

```

Prepping the data for visualizing

```{r}
library(tidyverse)
library(plotly)
library(forcats)
library(stringr)

# df_clean <- read_csv("tpbm_clean.csv")

studies <- df_clean %>%
  distinct(study_id, condition, .keep_all = TRUE)

pal_direction <- c(
  "positive"     = "#2a7f62",
  "mixed"        = "#e09b3d",
  "inconclusive" = "#7f8c8d",
  "negative"     = "#c0392b"
)
```

Stacked Bar Chart (conditions × outcome direction)

The stacked bar chart shows how many studies exist for each condition and how their results break down by outcome direction. Each horizontal bar represents a condition, and the total length reflects the total number of studies on that condition. The colored segments within each bar show how many of those studies reported positive, mixed, inconclusive, or negative results.
```{r}
d_bar <- studies %>%
  count(condition, outcome_direction) %>%
  mutate(condition = fct_reorder(condition, n, sum))

dirs <- c("positive", "mixed", "inconclusive", "negative")

p_bar <- plot_ly(type = "bar", orientation = "h")

for (dir in dirs) {
  sub <- d_bar %>% filter(outcome_direction == dir)
  if (nrow(sub) == 0) next
  
  p_bar <- p_bar %>% add_trace(
    x = sub$n,
    y = as.character(sub$condition),
    name = str_to_title(dir),
    marker = list(color = pal_direction[[dir]]),
    text = paste0(str_to_title(dir), ": ", sub$n),
    hoverinfo = "text+y"
  )
}

p_bar <- p_bar %>%
  layout(
    barmode = "stack",
    xaxis = list(title = "Number of studies"),
    yaxis = list(title = "")
  )

p_bar
```


Sankey Diagram... instead of alluvial. (study design to outcome)

The sankey diagram shows how study design types connect to outcome directions, highlighting where positive or negative results are coming from. The left side lists study designs (like RCTs or observational studies), and the right side shows outcome categories. The flows between them represent the number of studies, with thicker flows meaning more studies. Colors correspond to the outcome direction. You read it by following the flows from each design to see whether more rigorous designs tend to produce positive, mixed, or negative findings.
```{r}
d_sankey <- studies %>%
  filter(!is.na(outcome_direction)) %>%
  mutate(
    study_design = fct_lump_n(study_design, n = 5),
    study_design = as.character(study_design)
  ) %>%
  count(study_design, outcome_direction)

designs    <- sort(unique(d_sankey$study_design))
directions <- sort(unique(d_sankey$outcome_direction))

node_labels <- c(designs, directions)
n_designs   <- length(designs)

source_idx <- match(d_sankey$study_design, node_labels) - 1
target_idx <- match(d_sankey$outcome_direction, node_labels) - 1

dir_colors <- pal_direction

node_colors <- c(
  rep("#6a93d4", n_designs),
  sapply(directions, function(x) dir_colors[[x]])
)

link_colors <- sapply(d_sankey$outcome_direction, function(x) {
  col <- dir_colors[[x]]
  r <- strtoi(substr(col, 2, 3), 16)
  g <- strtoi(substr(col, 4, 5), 16)
  b <- strtoi(substr(col, 6, 7), 16)
  paste0("rgba(", r, ",", g, ",", b, ",0.4)")
})

p_sankey <- plot_ly(
  type = "sankey",
  node = list(
    label = node_labels,
    color = node_colors,
    pad = 20,
    thickness = 24
  ),
  link = list(
    source = source_idx,
    target = target_idx,
    value  = d_sankey$n,
    color  = link_colors
  )
)

p_sankey
```

Box plot — continuous session count distribution broken down by outcome direction, asks whether more sessions = better results
```{r}
d_box <- studies %>%
  filter(!is.na(session_count)) %>%
  mutate(
    direction_label = str_to_title(outcome_direction),
    tip = paste0(
      "<b>", author, " (", year, ")</b><br>",
      condition, "<br>",
      "Sessions: ", session_count, "<br>",
      str_to_title(outcome_direction)
    )
  )

dirs <- c("Positive", "Mixed", "Inconclusive", "Negative")
p_box <- plot_ly(type = "box")

for (dir in dirs) {
  sub <- d_box %>% filter(direction_label == dir)
  if (nrow(sub) == 0) next
  p_box <- p_box %>% add_trace(
    y         = sub$session_count,
    name      = dir,
    type      = "box",
    boxpoints = "all",
    jitter    = 0.4,
    pointpos  = 0,
    marker    = list(
      color   = pal_direction[[tolower(dir)]],
      size    = 5,
      opacity = 0.7
    ),
    line  = list(color = pal_direction[[tolower(dir)]]),
    fillcolor = paste0(
      "rgba(",
      paste(c(
        strtoi(substr(pal_direction[[tolower(dir)]], 2, 3), 16),
        strtoi(substr(pal_direction[[tolower(dir)]], 4, 5), 16),
        strtoi(substr(pal_direction[[tolower(dir)]], 6, 7), 16)
      ), collapse = ","),
      ",0.2)"
    ),
    text      = sub$tip,
    hoverinfo = "text"
  )
}

p_box <- p_box %>%
  layout(
    xaxis  = list(title = "Outcome direction"),
    yaxis  = list(title = "Number of sessions", zeroline = FALSE),
    legend = list(orientation = "h", x = 0, y = -0.15),
    showlegend = FALSE
  )

p_box
```

Sankey 2-  three-stage flow showing how pediatric, adult, and geriatric populations map to conditions and then to study designs
```{r}
d_sankey2 <- studies %>%
  filter(!is.na(population_age), !is.na(condition), !is.na(study_design)) %>%
  mutate(study_design = fct_lump_n(study_design, n = 6)) %>%
  count(population_age, condition, study_design) %>%
  filter(n >= 1)

ages     <- sort(unique(d_sankey2$population_age))
conds    <- sort(unique(d_sankey2$condition))
designs  <- sort(unique(as.character(d_sankey2$study_design)))

node_labels <- c(ages, conds, designs)
n_ages      <- length(ages)
n_conds     <- length(conds)

age_colors  <- c("adult" = "#2a5298", "geriatric" = "#8e44ad", "pediatric" = "#e09b3d")
cond_color  <- "#6a93d4"
des_color   <- "#95a5a6"

node_colors <- c(
  sapply(ages, function(x) age_colors[[x]]),
  rep(cond_color, n_conds),
  rep(des_color, length(designs))
)

source1 <- match(d_sankey2$population_age, node_labels) - 1
target1 <- match(d_sankey2$condition,      node_labels) - 1
source2 <- match(d_sankey2$condition,      node_labels) - 1
target2 <- match(as.character(d_sankey2$study_design), node_labels) - 1

p_sankey2 <- plot_ly(
  type = "sankey",
  node = list(
    label     = node_labels,
    color     = node_colors,
    pad       = 15,
    thickness = 20,
    line      = list(color = "#ffffff", width = 0.5)
  ),
  link = list(
    source = c(source1, source2),
    target = c(target1, target2),
    value  = c(d_sankey2$n, d_sankey2$n),
    label  = c(
      paste0(d_sankey2$population_age, " → ", d_sankey2$condition, ": ", d_sankey2$n),
      paste0(d_sankey2$condition, " → ", d_sankey2$study_design, ": ", d_sankey2$n)
    )
  )
)

p_sankey2 <- p_sankey2 %>%
  layout(margin = list(l = 20, r = 20, t = 30, b = 20))

p_sankey2

```

Bar chart 
```{r}
d_wl <- studies %>%
  filter(!is.na(wavelength_nm)) %>%
  mutate(wavelength_nm = as.factor(wavelength_nm)) %>%
  count(wavelength_nm, sort = TRUE)

p_wl <- plot_ly(
  d_wl,
  x         = ~fct_reorder(wavelength_nm, n),
  y         = ~n,
  type      = "bar",
  marker    = list(
    color = "#2a5298",
    line  = list(color = "#ffffff", width = 0.5)
  ),
  text      = ~paste0(wavelength_nm, " nm<br>", n, " studies"),
  hoverinfo = "text"
) %>%
  layout(
    xaxis = list(title = "Wavelength (nm)", tickangle = -45),
    yaxis = list(title = "Number of studies", zeroline = FALSE)
  )

p_wl

```

pie chart- study types
```{r}
d_pie <- studies %>%
  mutate(study_design = fct_lump_n(study_design, n = 6)) %>%
  count(study_design, sort = TRUE)

p_pie <- plot_ly(
  d_pie,
  labels    = ~study_design,
  values    = ~n,
  type      = "pie",
  textinfo  = "label+percent",
  hoverinfo = "label+value",
  marker    = list(
    colors = c("#2a5298", "#6a93d4", "#2a7f62", "#e09b3d", "#c0392b", "#8e44ad", "#95a5a6"),
    line   = list(color = "#ffffff", width = 1.5)
  )
) %>%
  layout(
    showlegend = TRUE,
    legend     = list(orientation = "h", x = 0, y = -0.15)
  )

p_pie

```

```{r}
d_box2 <- studies %>%
  filter(!is.na(session_duration_min), !is.na(population_type)) %>%
  mutate(
    population_type = str_to_title(population_type),
    tip = paste0(
      "<b>", author, " (", year, ")</b><br>",
      condition, "<br>",
      "Session duration: ", session_duration_min, " min<br>",
      str_to_title(population_type)
    )
  )

pop_pal <- c("Clinical" = "#2a5298", "Healthy" = "#2a7f62", "Aging" = "#8e44ad")

p_box2 <- plot_ly(type = "box")

for (pop in names(pop_pal)) {
  sub <- d_box2 %>% filter(population_type == pop)
  if (nrow(sub) == 0) next
  p_box2 <- p_box2 %>% add_trace(
    y         = sub$session_duration_min,
    name      = pop,
    type      = "box",
    boxpoints = "all",
    jitter    = 0.4,
    pointpos  = 0,
    marker    = list(
      color   = pop_pal[[pop]],
      size    = 5,
      opacity = 0.6
    ),
    line      = list(color = pop_pal[[pop]]),
    fillcolor = paste0(
      "rgba(",
      paste(c(
        strtoi(substr(pop_pal[[pop]], 2, 3), 16),
        strtoi(substr(pop_pal[[pop]], 4, 5), 16),
        strtoi(substr(pop_pal[[pop]], 6, 7), 16)
      ), collapse = ","),
      ",0.2)"
    ),
    text      = sub$tip,
    hoverinfo = "text"
  )
}

p_box2 <- p_box2 %>%
  layout(
    xaxis      = list(title = "Population type"),
    yaxis      = list(title = "Session duration (min)", zeroline = FALSE),
    showlegend = FALSE
  )

p_box2
```


